Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period

### 2019-12-14 analysis/paper/paper.Rmd
updates on the linear model

```{r chi-square-testing}
# prepare contingency table for chi sqaure
ornaments_shape_chi <-
  ornaments_period_clean %>% 
  group_by(period, Categories) %>% 
  tally() %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>% 
  spread(Categories, n) 
# compute chi sqaure
chisq_test_result <- chisq.test(ornaments_shape_chi[ , -1])
chisq_test_result_stat <- round(chisq_test_result$statistic, 2)
chisq_test_result_df <- round(chisq_test_result$parameter, 2)
chisq_test_result_p_value <- (chisq_test_result$p.value)
```

@@

Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequencies of the major classes of ornaments for different time periods at Kiwulan. The difference in frequencies between the three time periods reflect significant differences in the use of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = $`r chisq_test_result_p_value`$). Most ornament types were present before European contact. Ornament frequencies reached a peak during the European period and then dropped during the Chinese period, especially gold-foil beads. This trend can be also seen on other ornaments including carnelian beads, metal rings, and bells. However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European period and a further decrease in the Chinese period.

@@

How might these results fit into a bigger picture of social change at periphery of colonial systems? We may get some insight into the general pathways that led to social inequality in northeastern Taiwan by considering how people have achieved and maintained power in a wide variety of societies [@Drennan2010; @Feinman2000; @Ames2010; @Bowles2010]. The corporate/network model proposed by @Feinman2000 expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power. In the network mode, inequality develops when individuals accumulate wealth through their individual networks and people use their wealth to attract factions, control resources, and monopolize trade networks. In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999]. We may be able to interpreate Yilan social organisation as moving from corporate mode before the European arrival, to a network mode during European presence. The small number of ornaments, and less concentrated distribution during the pre-European period appears consistent with shared power and wealth of the corporate mode. The long-distance trade network introduced by Europeans resulted in the appearance of a network mode, and the emergence of competition among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994]. Because of the weak direct control from the European colonizers in northeastern Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].

>>>>

```{r chi-square-testing}
# prepare contingency table for chi sqaure
ornaments_shape_chi <-
  ornaments_period_clean %>% 
  group_by(period, Categories) %>% 
  tally() %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>% 
  spread(Categories, n) 
# compute chi sqaure
chisq_test_result <- chisq.test(ornaments_shape_chi[ , -1])
chisq_test_result_stat <- round(chisq_test_result$statistic, 2)
chisq_test_result_df <- round(chisq_test_result$parameter, 2)
chisq_test_result_p_value <- round(chisq_test_result$p.value,
                                   abs(floor(log10(chisq_test_result$p.value)-2)))
```

@@

```{r poisson-linear-model}
ornaments_potsherds <- read.csv(here("analysis", "data", "raw_data", "kwl_Ornament_potsherds.csv"))
## test: can we predict the number of ornaments using the mass of pottery?
# yes, seems like it

# poisson model
fit <- (glm(formula = sum_orna ~ sum_pot,
            family = "poisson",
            data = ornaments_potsherds))

linear_pvalue <- round(coef(summary(fit))[,'Pr(>|z|)'][2], 
                       abs(floor(log10(coef(summary(fit))[,'Pr(>|z|)'][2])-2)))

```

Figure \@ref(fig:plot-counts-by-period) shows the comparison of frequencies of the major classes of ornaments for different time periods at Kiwulan. The difference in frequencies between the three time periods reflect significant differences in the use of ornaments (chi-square = `r chisq_test_result_stat`, df = `r chisq_test_result_df`, p-value = $`r chisq_test_result_p_value`$). Most ornament types were present before European contact. Ornament frequencies reached a peak during the European period and then dropped during the Chinese period, especially gold-foil beads. This trend can be also seen on other ornaments including carnelian beads, metal rings, and bells. However, glass beads show a different pattern that indicates a higher frequency in the pre-European contact, and then a decrease in the European period and a further decrease in the Chinese period. When comparing with locally made pottery that could reflect Kiwulan population, the proportion of ornaments relative to the total mass of pottery remains the same, showing the general decreasing pattern in the Chinese period (Poisson regression model, p-value = $`r linear_pvalue`$).

@@

How might these results fit into a bigger picture of social change at periphery of colonial systems? We may get some insight into the general pathways that led to social inequality in northeastern Taiwan by considering how people have achieved and maintained power in a wide variety of societies [@Drennan2010; @Feinman2000; @Ames2010; @Bowles2010]. The corporate/network model proposed by @Feinman2000 expands traditional hierarchical complexity to provide a comparative basis for distinct strategies for power. In the network mode, inequality develops when individuals accumulate wealth through their individual networks and people use their wealth to attract factions, control resources, and monopolize trade networks. In contrast, the corporate mode stresses shared power across different groups and sectors, integrative ceremonies and rituals, and large cooperative labour tasks [@Feinman2000; @Siegel1999]. We may be able to interprete Yilan social organisation as moving from corporate mode before the European arrival, to a network mode during European presence. The small number of ornaments, and less concentrated distribution during the pre-European period appears consistent with shared power and wealth of the corporate mode. The long-distance trade network introduced by Europeans resulted in the appearance of a network mode, and the emergence of competition among ambitious individuals for prestige, wealth, or power through collecting trade goods [@Brumfiel1994; @Clark1994]. Because of the weak direct control from the European colonizers in northeastern Taiwan, local leaders may have had the flexibility to manipulate European colonial images, expand personal power, and monopolize the high-value trade goods [@Kang2012].
